Spatiotemporal models of `r paste(params$species)` biomass with climate covariates

### 2019-12-11 analysis/VOCC/2-biomass-depth-only-model.Rmd
update for reml version of sdmTMB; add year_cutoff for bioclimatic calculations

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Petrale Sole"
    region: "All synoptic surveys"
    knots: 600
    covariates: "+trawled+as.factor(ssid)" # additional fixed-effect covariates
    covs: "-trawled-ssid" # covariate string used to name model rds 
    update_model: TRUE
---

@@

  adult_biomass <- sdmTMB::sdmTMB(data,
    adult_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    ar1_fields = TRUE,
    include_spatial = TRUE,
    enable_priors = priors,
    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

@@

  imm_biomass <- sdmTMB::sdmTMB(data,
    imm_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
    include_spatial = TRUE,
    enable_priors = priors,
    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

@@

  total_biomass <- sdmTMB::sdmTMB(data,
    dens_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    ar1_fields = TRUE,
    include_spatial = TRUE,
    enable_priors = priors,
    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

>>>>

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    knots: 600
    covariates: "+trawled+as.factor(ssid) + trawled:as.factor(ssid)" # additional fixed-effect covariates
    covs: "-trawled-ssid-reml" # covariate string used to name model rds 
    update_model: TRUE
---

@@

  adult_biomass <- sdmTMB::sdmTMB(data=data,
    adult_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    #ar1_fields = TRUE,
    include_spatial = TRUE,
    reml = TRUE,
    enable_priors = priors,
    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

@@

  imm_biomass <- sdmTMB::sdmTMB(data=data,
    imm_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    #ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
    reml = TRUE,
    include_spatial = TRUE,
    enable_priors = priors,
    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

@@

  total_biomass <- sdmTMB::sdmTMB(data=data,
    dens_formula,
    time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
    time = "year", spde = spde,
    family = tweedie(link = "log"),
    ar1_fields = TRUE,
    include_spatial = TRUE,
    enable_priors = priors,
    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    silent = FALSE
  )

### 2019-12-11 analysis/VOCC/3-climate-models-fixed-both.Rmd
update for reml version of sdmTMB; add year_cutoff for bioclimatic calculations

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with climate covariates"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Pacific Ocean Perch"
    region: "All synoptic surveys"
    #region: "Both odd year surveys"
    log_temp: TRUE # should temperature be log()
    model_w_do: TRUE
    priors: FALSE
    knots: 500
    covariates: "+trawled" # additional non-climate variables
    covs: "-log-both-trawled" # string describing model covariates
---

@@

```{r params}
species <- params$species
region <- params$region
covariates <- params$covariates 
covs <- params$covs
log_temp <- params$log_temp
model_w_do <- params$model_w_do
priors <- params$priors
knots <- params$knots

paste("region =", region)
paste("model covariates =", covariates)
paste("model label =", covs)
paste("log(temperature) =", log_temp)
paste("model_w_do =", model_w_do)
paste("priors =", priors)
paste("knots =", knots)

@@

if (any(names(data) == "adult_density")) { 
  data <- data %>% select(X, Y, X10, Y10,
      adult_density, imm_density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, temp, temperature_c) 
 } else {
    data <- data %>% select(X, Y, X10, Y10,
      density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, temp, temperature_c) 
 }
}
```

@@

    adult_biomass_d <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    imm_biomass_d <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""))

    total_biomass_d <- sdmTMB::sdmTMB(data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
tictoc::toc()
```

>>>>

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with climate covariates"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    #region: "Both odd year surveys"
    log_temp: TRUE # should temperature be log()
    model_w_do: TRUE
    priors: FALSE
    knots: 800
    covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
    # covs: "-tv-depth-past" #
    # year_cutoff: 2013
    covs: "-tv-depth-all" #"-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
    year_cutoff: NULL
    #covs: "-log-both-noAR1" #"-log-both-noAR1" # string describing model covariates
---

@@

```{r params}
species <- params$species
region <- params$region
covariates <- params$covariates 
covs <- params$covs
log_temp <- params$log_temp
model_w_do <- params$model_w_do
priors <- params$priors
knots <- params$knots
year_cutoff <- params$year_cutoff

paste("region =", region)
paste("model covariates =", covariates)
paste("model label =", covs)
paste("log(temperature) =", log_temp)
paste("model_w_do =", model_w_do)
paste("priors =", priors)
paste("knots =", knots)
paste("year cutoff <", year_cutoff)

@@

if (any(names(data) == "adult_density")) { 
  data <- data %>% select(X, Y, X10, Y10,
      adult_density, imm_density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, temp, temperature_c) 
 } else {
    data <- data %>% select(X, Y, X10, Y10,
      density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, temp, temperature_c) 
 }
}

#   d <- data %>% filter(year==2011) #%>% filter(depth> 400) 
# = plot((adult_density)~(depth), data=d)

if (!is.null(year_cutoff)) data <- filter(data, year < year_cutoff)
```

@@

    adult_biomass_d <- sdmTMB::sdmTMB(data = data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      #ar1_fields = TRUE, 
      #include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp, covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    imm_biomass_d <- sdmTMB::sdmTMB(data = data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      #ar1_fields = TRUE,
      #include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp, covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    # dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""))
    # 
    # total_biomass_d <- sdmTMB::sdmTMB(data,
    #   dens_formula,
    #   time_varying = ~ 0 + depth_scaled + depth_scaled2, 
    #   time = "year", spde = spde,
    #   family = tweedie(link = "log"),
    #   #ar1_fields = TRUE,
    #   #include_spatial = FALSE, # default = TRUE
    #   enable_priors = priors,
    #   control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
    #   silent = FALSE
    # )
    # 
    # saveRDS(total_biomass_d, file = paste0("data/", spp, 
    #   "/model-total-biomass-", spp, covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
    #   ))
  }
tictoc::toc()
```

Save predictions for full spatial grid

```{r eval=FALSE}
#fish <- readRDS(paste0("raw/bio-data-", spp, ""))

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

if (any(names(data) == "adult_density")) {
#if (length(unique(fish$maturity_code)) > 2) {
  nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))

  adult_biomass <- readRDS(paste0("data/", spp, 
        "/mod-mat-biomass-", spp, covs,"-", year_cutoff, "-", ssid_string, 
        "-prior-", priors, ".rds"))

  nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(adult_biomass$data$year)) %>% 
    mutate(
      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
      do_mlpl_scaled2 = do_mlpl_scaled^2) %>% 
    select(-est, -est_non_rf, - est_rf, -omega_s, -epsilon_st, -zeta_s)
  nd <- na.omit(nd)
  #nd$year <- as.integer(nd$year)

  adult_predictions <- predict(adult_biomass, newdata = nd)
  saveRDS(adult_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs,"-", year_cutoff, "-", ssid_string, 
        "-prior-", priors, ".rds"))

  imm_biomass <- readRDS(paste0("data/", spp, 
        "/mod-imm-biomass-", spp, covs,"-", year_cutoff, "-", ssid_string, 
        "-prior-", priors, ".rds"))

    nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(imm_biomass$data$year)) %>% 
    mutate(
      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
      do_mlpl_scaled2 = do_mlpl_scaled^2) %>% 
    select(-est, -est_non_rf, - est_rf, -omega_s, -epsilon_st, -zeta_s)
  nd <- na.omit(nd)
  #nd$year <- as.integer(nd$year)

  imm_predictions <- predict(imm_biomass, newdata = nd)
  saveRDS(imm_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs,"-", year_cutoff, "-", ssid_string, 
        "-prior-", priors, ".rds"))

    # total_biomass <- readRDS(paste0("data/", spp,
    #   "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
    # 
    # nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
    # nd <- nd_all %>% 
    #   filter(ssid %in% model_ssid) %>%
    #   filter(year %in% unique(total_biomass$data$year))%>% 
    # mutate(
    #   do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
    #   do_mlpl_scaled2 = do_mlpl_scaled^2) %>% 
    # select(-est, -est_non_rf, - est_rf, -omega_s, -epsilon_st, -zeta_s)
    # nd <- na.omit(nd)
    # #nd$year <- as.integer(nd$year)
    # 
    # predictions <- predict(total_biomass, newdata = nd)
    # saveRDS(predictions, file = paste0("data/", spp,
    #   "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
}
```

### 2019-12-11 analysis/VOCC/3-climate-models-fixed-depth.Rmd
update for reml version of sdmTMB; add year_cutoff for bioclimatic calculations

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with climate covariates"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    #region: "Both odd year surveys"
    log_temp: TRUE # should temperature be log()
    model_w_do: TRUE
    priors: FALSE
    knots: 500
    covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
    covs: "-AR1-spatial" #
    #covs: "-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
---

@@

# d <- data %>% filter(year==2011) #%>% filter(depth> 400) 
# = plot((adult_density)~(depth), data=d)

```

@@

    adult_biomass_d <- sdmTMB(data = data,
      adult_formula,
      #time_varying = ~ trawled, #0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE, 
      include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    imm_biomass_d <- sdmTMB::sdmTMB(data = data, 
      imm_formula,
      #time_varying = ~ trawled, # 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
tictoc::toc()
```

>>>>

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with climate covariates"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    #species: "Redbanded Rockfish"
    #species: "Pacific Cod"
    #species: "Pacific Ocean Perch"
    #species: "Arrowtooth Flounder" #all DO smiley
    #species: "Petrale Sole" #adult DO almost quadratic...
    #species: "English Sole" #all DO smiley
    #species: "Dover Sole" #all DO smiley
    species: "Flathead Sole"
    region: "All synoptic surveys"
    #region: "Both odd year surveys"
    log_temp: TRUE # should temperature be log()
    model_w_do: TRUE
    priors: FALSE
    knots: 600
    # knots: 500
    covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
    covs: "-past" # hardcoded "-fixed" makes covs = "-fixed-past"
    year_cutoff: 2013
    # covs: "-all" # hardcoded -fixed makes covs = "-fixed-all"
    # year_cutoff: NULL
    #"-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
---

@@

# d <- data %>% filter(year==2011) #%>% filter(depth> 400) 
# = plot((adult_density)~(depth), data=d)
if (covs == "-past") data <- filter(data, year < year_cutoff)
```

@@

    adult_biomass_d <- sdmTMB(data = data,
      adult_formula,
      #time_varying = ~ trawled, #0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      #ar1_fields = TRUE, 
      #include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed", covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    imm_biomass_d <- sdmTMB::sdmTMB(data = data, 
      imm_formula,
      #time_varying = ~ trawled, # 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      #ar1_fields = TRUE,
      #include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed", covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
        ))

@@

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed", covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
tictoc::toc()
```
